Cooling Requirements for the Vertical Shear Instability in Protoplanetary Disks

### 2015-05-03 linear_adia.tex
very slight correction to the sentence before eq 47
finally resolved disagreement over sec 5.3, eq 48

Eq. \ref{relax_cond_simp1} admits the following   
approximate solution for the frequency if $M|hq\khat|^2/(2M+1)\lesssim
1$, 
\begin{align}
  % \left(\hat{\omega}_c\khat\right)^2 \simeq (M+1)\left[1  + O( M|h q \khat|^2)\right]  \label{marg_freq_approx}
  \left(\hat{\omega}_c\khat\right)^2 \simeq (M+1)\left(1 -
    \frac{M|h q \khat|^2}{2M+1}\right).  \label{marg_freq_approx}
\end{align}  
% which holds for either (a) $M= 0$ and arbitrary $|h q \hat{k}|$ or (b) $|h q \hat{k}| \ll 1$ and arbirtary $M \sim O(1)$. 
% These distinct conditions mean that our results are most robust for the fundamental vertical mode ($M = 0$) but also
% hold for higher order vertical modes when radial wavelengths are not
% too short. 
Note that Eq. \ref{marg_freq_approx} reduces to the dispersion
relation for low-frequency inertial waves in the absence of vertical
shear (see Appendix \ref{stable_novshear} for details).

If we further assume $|hq\khat|^2\ll 1$, we obtain from
Eq. \ref{relax_cond_simp2}---\ref{marg_freq_approx} that  
\begin{align}\label{bcrit_gen} 
  \beta_c &\simeq  \frac{h|q|}{(\gamma - 1)}\frac{\sqrt{1+M}}{(1+2M)}.   
%  \frac{\sqrt{1+M}}{(1+2M)}
%  \left[1 + \frac{M(2M-1)|h q   \khat|^2}{2(2M+1)^2}\right].
%  &\lesssim \beta_\mathrm{crit}%\notag
%>>>>>>> 8816838e4a06974fc61d74b84a2474adaeee3ebc
\end{align}
% but the correction term ($O(M |h q \khat|^2)$) is now suppressed.    
Eq. \ref{bcrit_gen} indicates that higher $M$ modes have lower $\beta_c$,
i.e.\ require faster cooling for VSI growth\footnote{In fact,   
  this conclusion does not require $|hq\khat|^2$ to be very 
  small. The may be seen from the more complete expression for
  $\beta_c$ obtained by substituting Eq. \ref{marg_freq_approx} into  
  Eq. \ref{relax_cond_simp2} and retaining the term proportional to 
  $|hq\khat|^2$.}.    
This trend is confirmed numerically in \S\ref{therm_relax_eff}.

@@

We are most interested in the longest cooling time which allows growth for any $M$.  
This criterion is clearly set by the fundamental, $M = 0$, mode.  Our
cooling criterion for  
VSI growth is thus  
 \begin{align}\label{iso_vsi_cond}
  \beta_c\leq \beta_\mathrm{crit}  \equiv
  \frac{h|q|}{\gamma-1} . 
\end{align}
The thin disk approximation, $h \ll 1$, indicates that $\beta_c \ll
1$, as assumed to obtain Eq. \ref{marg_freq_approx}. We heuristically
explain $\beta_\mathrm{crit}$ in \S\ref{vsi_require}, and numerically
test its validity in \S\ref{bcrit_num_test}.

>>>>

Eq. \ref{relax_cond_simp1} admits the following   
approximate solution for the frequency 
if $\mathrm{min}(1,M)\times|hq\khat|^2\lesssim 1$, 
%if $4M(M+1)|hq\khat|^2/(2M+1)^2 < 1$,  
\begin{align}
  % \left(\hat{\omega}_c\khat\right)^2 \simeq (M+1)\left[1  + O( M|h q \khat|^2)\right]  \label{marg_freq_approx}
  \left(\hat{\omega}_c\khat\right)^2 \simeq (M+1)\left(1 -
    \frac{M|h q \khat|^2}{2M+1}\right).  \label{marg_freq_approx}
\end{align}  
(Equality holds for $M=0$.)  
% which holds for either (a) $M= 0$ and arbitrary $|h q \hat{k}|$ or (b) $|h q \hat{k}| \ll 1$ and arbirtary $M \sim O(1)$. 
% These distinct conditions mean that our results are most robust for the fundamental vertical mode ($M = 0$) but also
% hold for higher order vertical modes when radial wavelengths are not
% too short. 
Note that Eq. \ref{marg_freq_approx} reduces to the dispersion
relation for low-frequency inertial waves in the absence of vertical
shear (see Appendix \ref{stable_novshear} for details).

If we further assume $|hq\khat|^2\ll 1$, we obtain from
Eq. \ref{relax_cond_simp2}---\ref{marg_freq_approx} that  
\begin{align}\label{bcrit_gen} 
  \beta_c &\simeq  \frac{h|q|}{(\gamma - 1)}\frac{\sqrt{1+M}}{(1+2M)}.   
%  \frac{\sqrt{1+M}}{(1+2M)}
%  \left[1 + \frac{M(2M-1)|h q   \khat|^2}{2(2M+1)^2}\right].
%  &\lesssim \beta_\mathrm{crit}%\notag
%>>>>>>> 8816838e4a06974fc61d74b84a2474adaeee3ebc
\end{align}
Eq. \ref{bcrit_gen} indicates that higher $M$ modes have lower $\beta_c$,
i.e.\ require faster cooling for VSI growth\footnote{
In fact, this conclusion does not require $\theta\equiv|hq\khat|^2$ to be extremely 
small. By retaining terms proportional to
  $\theta$, we find $\beta_c\to \beta_c \times 
  \left[1 + M(2M-1)\theta/2(2M+1)^2 \right]$. Since $\theta\lesssim 1$ by assumption, 
 $\mathrm{max}(\beta_c)$ still occurs at $M=0$.}.     
This trend is confirmed numerically in \S\ref{therm_relax_eff}.

@@

We are most interested in the longest cooling time which allows growth for any $M$.  
This criterion is clearly set by the fundamental, $M = 0$, mode.  Our
cooling criterion for  
VSI growth is thus  
 \begin{align}\label{iso_vsi_cond}
  \beta_c\leq \beta_\mathrm{crit}  \equiv
  \frac{h|q|}{\gamma-1} . 
\end{align}
The thin disk approximation, $h \ll 1$, indicates that $\beta_c \ll
1$, as assumed to obtain Eq. \ref{relax_cond_simp1}---\ref{relax_cond_simp2}. 
We heuristically explain $\beta_\mathrm{crit}$ in \S\ref{vsi_require}, and numerically
test its validity in \S\ref{bcrit_num_test}.
